Seed Bank Data Compilation
========================================================

During the summer of 2011 I manipulated 10 plots at UNH-West Foss Farm and
10 plots at LI-Caleb Smith in an attempt to examine affects of seedbank
dynamics on *Frangula alnus* (Glossy buckthorn) population dynamics.
The manipulation was as following:

* I randomly chose 10 lat/long points each, contained within the 
two previously established study areas.
* I went to each point using a Garmin GPS.
* At that points, I established a 1x1 meter plot in which the side
of the plot was parallel to the magnetic North-South direction.
* I cleared the plot of all plant material using a Council Fire Rake.
Larger plants were cut at their base. All plant material was discarded
nearby. 
* The cleared 1x1 m plot was seperated into four quadrats. Oppisite 
quadrats were selected at random and covered with landscaping cloth.

These plots were setup at UNH-West Foss Farm on 17 & 18 AUG 2011 and
at LI-Caleb Smith on 24 AUG 2011.

On 7 MAY 2012 and 24 MAY 2012 I returned to UNH-WFF and LI-CS, respectively,
and removed the landscape cloth. I counted all emerging seedlings in all 
of the quadrats. Seedlings were tagged with a small plastic tag to facilitate
counting at a later time. The collected data have mostly been transferred 
from my field notebooks into the Excel file:
`/Users/mlammens/Google Drive/F-alnus/Field-Work`

Here I am documenting the steps I am taking to make this dataset usable
in R for further analysis.

## Seedbank CSV File

I wrote a CSV file in the Chapter 5 Project folder, 
subfolder `/data` named `FRAL_seedbank_plots.csv` that 
contains the values location, site, plot, Aug. 2011 notes, and May 2012 notes. As originally written,
each plot is represented in a single row, but I want to seperate these
out into four rows per plot based on the quadrats.

R Code Off
```
# Read in csv
seedbank <- read.csv('data/FRAL_seedbank_plots.csv')
# Define a vector of quadrat names
quad.names <- c('NE','NW','SW','SE')
# Create a data.frame in which each row from seedbank is repeated
# four times.
seedbank.full <- seedbank[ rep(row.names(seedbank), each=4),1:3 ]
# Combine this data.frame with a column of the quadrat names
seedbank.full <- data.frame(seedbank.full,quadrat=rep(quad.names,20))
# Now write this new data.frame out as a csv file, replacing the 
# original csv file.
write.csv(seedbank.full,'data//FRAL_seedbank_seedlingCnt.csv',quote=FALSE,
          row.names=FALSE)
```

## Seedling Counts

Using the `seedbank.full` file, I manually set the quadrats that 
were covered or not, and the number of seedlings found in each quadrat
during survey periods. There were two survey periods for UNH-WFF and three
survey periods for LI-CS. 

For UNH-WFF, survey periods were:

* 7 May 2012: All covered quadrats were uncovered and any seedlings were
counted and tagged with a white plastic clip.
* 24 Oct 2012: All quadrats were examined to see if white plastic clips
were still present and to count the total number of seedlings in the quadrat.
**No new seedlings were tagged at this time**

For LI-CS, survey periods were:

* 24 May 2012: All covered quadrats were uncovered and any seedlings were
counted and tagged with a white plastic clip.
* 27 Jun 2012: All quadrats were examined for a second time. Some tags were
recovered (no longer attached, presumumbly because the seedling died). Any new
seedlings were tagged with a white plastic clip.
* __ Nov 2012: All quadrats were examined to see if white plastic clips
were still present and to count the total number of seedlings in the quadrat.
**No new seedlings were tagged at this time**

I added the above seedling counts to the CSV file manually, based on
field notes in my field notebooks and from field data sheets. My field
data sheets have also been scanned, and a digital copy is here:
`/Users/mlammens/Google Drive/F-alnus/Field-Work`

## Using the new dataset

### Change NAs to 0

Here I read-in and manipulate this data set. One important manipulation I do
here is to make all NA values equal to 0. This is to make calculations easier.
Technically, if value in the dataset is NA, it is because either I did not 
make a related observation or that variable was not collected at all. An
example of the latter is the NAs for the `seedlingCnt.jun.2012` values for 
UNH-WFF. I did not collect any such values from UNH-WFF, only from LI-CS.
However, for the purposes of calculating various metrics, it is far easier if
this values are 0s.

```{r}
# Define project directory
FRAL_DEMOG_DIR <- '/Users/mlammens/Google Drive/F-alnus/Chapter-5-FRAL-Demography/'
# Read seedbank_seedling csv
seedbank <- read.csv(paste(FRAL_DEMOG_DIR,'data/FRAL_seedbank_seedlingCnt.csv',sep=''))
# Make all NA values equal to 0.
# Be careful using `replace` here, because it could change
# factorial columns to numbers, or some other error. I checked 
# that it worked here.
seedbank <- replace(seedbank,is.na(seedbank),0)
```

### Calculate mean seedling count for covered vs. uncovered quadrats

```{r}
# Require plyr
require(plyr)

# Mean seedling count for covered vs. uncovered - May 2012
ddply(.data=seedbank,.variable=c('covered'),fun=mean(seedlingCnt.may.2012),summarize)

# Mean seedling count for covered vs. uncovered - Fall 2012
ddply(.data=seedbank,.variable=c('covered'),fun=mean(seedlingCnt.fall.2012),summarize)
```

### Calculate mean seedling count for covered vs. uncovered quadrats separately for WFF and CS

```{r}
# Mean seedling count for cov. vs. uncov. - May 2012 - WFF and CS separately
ddply(.data=seedbank,.variable=c('covered','site'),fun=mean(seedlingCnt.may.2012),summarize)

# Mean seedling count for cov. vs. uncov. - Fall 2012 - WFF and CS separately
ddply(.data=seedbank,.variable=c('covered','site'),fun=mean(seedlingCnt.fall.2012),summarize)
```

### Calculate the mortality rate for the seedbank plots

Basing this on the number of **tagged** seedlings where the tags were later recovered

```{r}
# divide the `may.to.fall.2012.mortality` by the sum of 
# `seedlingCnt.may.2012` and `seedlingCnt.jun.2012`
# Note that I'm removing the tags that were missing, assuming
# that I cannot classify them as dead or alive
seedling.mortality.plot <- seedbank$may.to.fall.2012.mortality /
  (seedbank$seedlingCnt.may.2012+seedbank$seedlingCnt.jun.2012-seedbank$may.to.fall.2012.missing)
# Calcualte the mean value
mean(seedling.mortality.plot,na.rm=TRUE)
```

